Changes in Patterns of Geometric Sign Groupings as a Model of Social Dynamics Across Aurignacian Phases

### 2025-04-14 endofquarterpaper.qmd
update bad citation

These results are consistent with previous work on the environmental and population change during these phases. From 45 - 43.25 k BP, the population density in Europe was extremely low, due to the unfavorable environmental conditions of Greenland Stadial 12 (44 k BP) and the slow westward expansion of early modern humans [@Shao_Wegener_Klein_Schmidt_Weniger_2024]. This was followed by rapid expansion from 43.25 - 41 k BP (Transitional/Proto), with the maximum extent of Proto/Early Aurignacian settlement being reached around 41 k BP. This rapid population expansion could be what stimulated the emergence of signs as a marker of group membership. Other studies have also found similar associations in other regions and time periods between increases in quantity and diversity of art, and increases in population density \[@smith1992Design; @hays1993symbol; @McDonald_Veth_Lilley_2006; @Lourandos_1985; @barton_clark_cohen1994; @mcdonald2012social; @bernardini2005reconsidering; @wiessner1984reconsidering; @smith1992\].

@@

>>>>

These results are consistent with previous work on the environmental and population change during these phases. From 45 - 43.25 k BP, the population density in Europe was extremely low, due to the unfavorable environmental conditions of Greenland Stadial 12 (44 k BP) and the slow westward expansion of early modern humans [@Shao_Wegener_Klein_Schmidt_Weniger_2024]. This was followed by rapid expansion from 43.25 - 41 k BP (Transitional/Proto), with the maximum extent of Proto/Early Aurignacian settlement being reached around 41 k BP. This rapid population expansion could be what stimulated the emergence of signs as a marker of group membership. Other studies have also found similar associations in other regions and time periods between increases in quantity and diversity of art, and increases in population density [@smith1992Design; @hays1993symbol; @McDonald_Veth_Lilley_2006; @Lourandos_1985; @barton_clark_cohen1994; @mcdonald2012social; @bernardini2005reconsidering; @wiessner1984reconsidering; @smith1992].

@@

<-- cite signbase 2.0 -->.
